# Analysis: Unicox
## Unicox:GEO
```{r}
dir.create("04.rmps_cluster")
geo = readRDS("01.datatidy/GEO_afterBatch.rds")
rmps = readRDS("01.datatidy/rmps.rds")

## mapping data
rmps$Symbol[!rmps$Symbol %in% row.names(geo$exp)]
row.names(geo$exp)[which(row.names(geo$exp) == "BUD23")] = "WBSCR22"
row.names(geo$exp)[which(row.names(geo$exp) == "MRM2")] = "FTSJ2"
row.names(geo$exp)[which(row.names(geo$exp) == "MRM3") ]= "RNMTL1"
rmps$Symbol[!rmps$Symbol %in% row.names(geo$exp)]

phe = geo$phe
sur = data.frame(os = phe$os.event, os.time = phe$os.time, id = phe$GSM) %>% 
  na.omit() %>% 
  subset(id %in% colnames(geo$exp))

exp = geo$exp[intersect(rmps$Symbol, row.names(geo$exp)), sur$id] %>%
  t() %>%
  data.frame()

## unicox test
library(survival)
library(survminer)
subExp = as.list(exp)

uniCoxT = function(dat){
  dat = data.frame(os = sur$os, os.time = sur$os.time, 
                 gene = dat)
  res.cox <- coxph(Surv(os.time, os) ~ gene, data = dat)
  res.cox <- summary(res.cox)
  p.value<-signif(res.cox$wald["pvalue"], digits=3)
  HR <-signif(res.cox$coef[2], digits=3)
  HR.confint.lower <- signif(res.cox$conf.int[,"lower .95"], 3)
  HR.confint.upper <- signif(res.cox$conf.int[,"upper .95"], 3)
  out = data.frame(pval = p.value, HR = HR, HR.up = HR.confint.upper,
                   HR.low = HR.confint.lower)
  return(out)
}

uni.res = lapply(subExp, uniCoxT)
uni.res = do.call(rbind, uni.res)
saveRDS(uni.res, file = "04.rmps_cluster/geo_uni.rds")

uni.res$gene = row.names(uni.res)
write.table(uni.res, file = "04.rmps_cluster/geo_unicox.xls", sep = "\t", row.names = F)

# visualization
uni.res$gene = fct_rev(uni.res$gene)
uni.res$type = ifelse(uni.res$HR>1, "bad", "good")
uni.res$type = ifelse(uni.res$pval<0.05, uni.res$type, "non")

ggplot(dat = uni.res, aes(y = gene, color = type)) + 
  theme_classic() +
  geom_point(aes(x=HR), shape=15, size=3)+
  geom_linerange(aes(xmin=HR.low, xmax=HR.up), linewidth = 1) +
  geom_vline(xintercept = 1, linetype="dashed") +
  labs(x="Hazard Ratio", y="") +
  coord_cartesian(xlim=c(0.5, 1.5))+ 
  theme(axis.line.y = element_blank(),
        axis.ticks.y= element_blank(),
        axis.text.y= element_blank(),
        axis.title.y= element_blank()) +
  scale_color_manual(values = c(`good` = "#238b45",
                                `bad` = "#e41a1c",
                                `non` = "#737373")) +
  theme(legend.position = "none")-> p_mid

uni.res$label = paste(uni.res$HR, "(", uni.res$HR.low, "-", uni.res$HR.up, ")", sep = "")
ggplot(dat = uni.res, aes(y = gene)) +
  geom_text(aes(x = 0, label = gene), hjust = 0, fontface = "italic")+
  geom_text(
    aes(x = 1, label = label),
    hjust = 0,
    fontface = ifelse(uni.res$label == "Hazard Ratio (95% CI)", "bold", "plain")
  )+
  theme_void() +
  coord_cartesian(xlim = c(0, 4)) -> p_left


ggplot(dat = uni.res) +
geom_text(
  aes(x = 0, y = gene, label = pval),
  hjust = 0,
  fontface = ifelse(uni.res$pval == "p-value", "bold", "plain")
) +
theme_void() -> p_right

library(patchwork)
layout <- c(
  area(t = 0, l = 0, b = 47, r = 3), 
  area(t = 0, l = 4, b = 47, r = 4), 
  area(t = 0, l = 5, b = 47, r = 9) 
)
# final plot arrangement
plot(layout)
out = p_left + p_mid + p_right + plot_layout(design = layout)
ggsave(out, filename = "04.rmps_cluster/unicox_geo.pdf", width = 12, height = 7)
```

## Unicox:TCGA
```{r}
rmpExp = readRDS("01.datatidy/RMP_exp.rds")
tcga_phe = readRDS("01.datatidy/tcga_phe.rds")

library(dplyr)
crcPhe = tcga_phe$phe %>% 
  subset(program == "TCGA") %>% 
  subset(project_id %in% c("TCGA-COAD", "TCGA-READ")) %>%
  subset(stringr::str_length(sample) == 16)  %>%
  subset(stringr::str_sub(sample, 14, 16) == "01A") %>%
  mutate(id = stringr::str_sub(sample, 1, 12)) 

crcSur = tcga_phe$sur %>% 
  subset(sample %in% crcPhe$sample) %>%
  left_join(crcPhe, by = "sample") %>%
  subset(sample %in% row.names(rmpExp))

crcExp = rmpExp[crcSur$sample,] %>% data.frame()

subExp = as.list(crcExp)

uniCoxT = function(dat){
  dat = data.frame(os = crcSur$OS, os.time = crcSur$OS.time, 
                 gene = dat)
  res.cox <- coxph(Surv(os.time, os) ~ gene, data = dat)
  res.cox <- summary(res.cox)
  p.value<-signif(res.cox$wald["pvalue"], digits=3)
  HR <-signif(res.cox$coef[2], digits=3)
  HR.confint.lower <- signif(res.cox$conf.int[,"lower .95"], 3)
  HR.confint.upper <- signif(res.cox$conf.int[,"upper .95"], 3)
  out = data.frame(pval = p.value, HR = HR, HR.up = HR.confint.upper,
                   HR.low = HR.confint.lower)
  return(out)
}

uni.res = lapply(subExp, uniCoxT)
uni.res = do.call(rbind, uni.res)
saveRDS(uni.res, file = "04.rmps_cluster/tcga_uni.rds")

uni.res$gene = row.names(uni.res)
write.table(uni.res, file = "04.rmps_cluster/tcga_unicox.xls", sep = "\t", row.names = F)

uni.res$gene = fct_rev(uni.res$gene)
uni.res$type = ifelse(uni.res$HR>1, "bad", "good")
uni.res$type = ifelse(uni.res$pval<0.05, uni.res$type, "non")

ggplot(dat = uni.res, aes(y = gene, color = type)) + 
  theme_classic() +
  geom_point(aes(x=HR), shape=15, size=3)+
  geom_linerange(aes(xmin=HR.low, xmax=HR.up), linewidth = 1) +
  geom_vline(xintercept = 1, linetype="dashed") +
  labs(x="Hazard Ratio", y="") +
  coord_cartesian(xlim=c(0.5, 1.5))+ 
  theme(axis.line.y = element_blank(),
        axis.ticks.y= element_blank(),
        axis.text.y= element_blank(),
        axis.title.y= element_blank()) +
  scale_color_manual(values = c(`good` = "#238b45",
                                `bad` = "#e41a1c",
                                `non` = "#737373")) +
  theme(legend.position = "none")-> p_mid

uni.res$label = paste(uni.res$HR, "(", uni.res$HR.low, "-", uni.res$HR.up, ")", sep = "")
ggplot(dat = uni.res, aes(y = gene)) +
  geom_text(aes(x = 0, label = gene), hjust = 0, fontface = "italic")+
  geom_text(
    aes(x = 1, label = label),
    hjust = 0,
    fontface = ifelse(uni.res$label == "Hazard Ratio (95% CI)", "bold", "plain")
  )+
  theme_void() +
  coord_cartesian(xlim = c(0, 4)) -> p_left


ggplot(dat = uni.res) +
geom_text(
  aes(x = 0, y = gene, label = pval),
  hjust = 0,
  fontface = ifelse(uni.res$pval == "p-value", "bold", "plain")
) +
theme_void() -> p_right

library(patchwork)
layout <- c(
  area(t = 0, l = 0, b = 47, r = 3), 
  area(t = 0, l = 4, b = 47, r = 4), 
  area(t = 0, l = 5, b = 47, r = 9) 
)
# final plot arrangement
plot(layout)
out = p_left + p_mid + p_right + plot_layout(design = layout)
ggsave(out, filename = "04.rmps_cluster/unicox_tcga.pdf", width = 12, height = 7)
```

# Analysis: Consensus Cluster
```{r}
geo = readRDS("01.datatidy/GEO_afterBatch.rds")
rmps = readRDS("01.datatidy/rmps.rds")

# mapping gene
rmps$Symbol[which(!rmps$Symbol %in% row.names(geo$exp))]
row.names(geo$exp)[which(row.names(geo$exp) == "MRM2")] = "FTSJ2"
row.names(geo$exp)[which(row.names(geo$exp) == "BUD23")] = "WBSCR22"
row.names(geo$exp)[which(row.names(geo$exp) == "MRM3") ]= "RNMTL1"
table(!rmps$Symbol %in% row.names(geo$exp))

# consensus cluster analysis
library(ConsensusClusterPlus)

dat = as.data.frame(geo$exp)
dat = dat[rmps$Symbol,]
dat = na.omit(dat)
dat = sweep(dat,1, apply(dat,1,median,na.rm=T))

title="04.rmps_cluster/cluster"
results = ConsensusClusterPlus(as.matrix(dat),maxK=6,reps=50,pItem=0.8,pFeature=1,
                               title=title,clusterAlg="hc",
                               distance="pearson",seed=12345678,
                               plot="png")
saveRDS(results, file = "04.rmps_cluster/cluster.rds")

# visualization
dat = as.data.frame(geo$exp)
dat = dat[rmps$Symbol,]
dat = na.omit(dat)
# dat = sweep(dat,1, apply(dat,1,median,na.rm=T))

phe = geo$phe
columnAnn = data.frame(id = colnames(dat), 
                       clusters = as.character(results[[3]]$consensusClass),
                       datasets = phe$GSE[match(colnames(dat), phe$GSM)])
columnAnn = columnAnn[order(columnAnn$clusters),]

library(ComplexHeatmap)
library(RColorBrewer)

# define colors
# col_fun = c("#5e3c99", "#b2abd2", "#f7f7f7", "#fdb863", "#e66101")
col_fun = brewer.pal(5, "PuOr")
col_fun = colorRampPalette(rev(col_fun))(100)

datasets_col = colorRampPalette(brewer.pal(10, "Paired"))(12)
names(datasets_col) = unique(columnAnn$datasets)

# annotation
column_ha = HeatmapAnnotation(Clusters = columnAnn$clusters,
                              Datasets = columnAnn$datasets,
               col = list(Clusters = c("1" = "#ec7014", "2" = "#88419d"
                                       ,
                                       "3" = "#238b45"
                                       ),
                          Datasets = datasets_col))
modType = rmps$ModificationSymbol
# modType[which(modType == "m1G:m1A") ] = "m1G"
# modType[which(modType == "m2,2,7G")] = "m2G"
# modType[which(modType == "m2,2G")] = "m2G"
# modType[which(modType == "m5c")] = "m5C"
# modType[which(modType == "m6A(?)")] = "m6A"
# modType[which(modType == "m6A(m)(Cap)")] = "m6A"
# modType[which(modType == "Xm(3'end)")] = "Xm"

row_ha = rowAnnotation(ModType = modType)

dat = dat[,columnAnn$id]

scale_data = function(x){
  # x = (x-mean(x))/sd(x)
  # x = (x-min(x))/(max(x) - min(x))
  x = x - mean(x)
  x = x / max(abs(x))
  return(x)
}
dat = apply(dat, 1, scale_data)
dat = t(dat)

pdf(file = "04.rmps_cluster/rmpCluster.pdf", width = 8, height = 8.5)
Heatmap(dat, col = col_fun,
        clustering_distance_columns  = "pearson",
        clustering_distance_rows  = "pearson",
        show_column_dend = FALSE,
        show_row_dend = FALSE,
        column_order = columnAnn$id,
        show_column_names  = FALSE,
        row_names_gp = gpar(fontface = "italic"),
        column_split = columnAnn$clusters,
        row_split = modType,
        top_annotation = column_ha,
        left_annotation = row_ha)
dev.off()
```

# Analysis: Cluster & Survival
```{r}
cluster = readRDS("04.rmps_cluster/cluster.rds")
geo = readRDS("01.datatidy/GEO_afterBatch.rds")
phe = geo$phe
cluster = cluster[[3]]$consensusClass
phe = geo$phe
dat = data.frame(os = phe$os.event, os.time = phe$os.time, cluster = cluster, GSE = phe$GSE)
dat = subset(dat, GSE == "GSE39582") # GSE39582, GSE17536
dat = na.omit(dat)

## survival analysis---------
library(dplyr)
library(survival)
library(survminer)
sfit <- survfit(Surv(os.time, os)~cluster, 
                data=dat)

pdf(file = "04.rmps_cluster/rmpSurvial_GSE39582.pdf", width = 4, height = 5.5)
ggsurvplot(sfit, 
           pval=TRUE,
           palette = c("#ec7014", "#88419d","#238b45"),
           risk.table = TRUE)
dev.off()
```
# Analysis: Cluster & MMR
```{r}
cluster = readRDS("04.rmps_cluster/cluster.rds")
geo = readRDS("01.datatidy/GEO_afterBatch.rds")
phe = geo$phe
cluster = cluster[[3]]$consensusClass

phe = geo$phe
dat = data.frame(mmr = phe$mmr.status, cluster = as.character(cluster))
dat = na.omit(dat)
dat = subset(dat, mmr %in% c("dMMR",  "MSI",  "MSS", "pMMR"))

library(ggplot2)
dat$Cluster = paste("cluster", dat$cluster, sep = "")
dat = data.frame(table(dat))
dat = split(dat, dat$mmr)
dat = lapply(dat, function(x){x$Percent = (x$Freq/sum(x$Freq))*100
return(x)})
dat = do.call(rbind, dat)
dat$Percent = round(dat$Percent, 2)
dat$mmr = factor(dat$mmr, levels = c("dMMR", "pMMR", "MSI", "MSS"))

pdf(file = "04.rmps_cluster/rmps_MMR.pdf", width = 2, height = 3.5)
ggplot(dat, aes(x = mmr, y = Percent, fill=Cluster)) +
  geom_bar(position='stack', stat='identity') + 
  theme_classic() + 
  theme(panel.grid=element_blank()) +
  ylab("Precent(100%)") +
  scale_fill_manual(values = c("cluster1" = "#ec7014",
                               "cluster2" = "#88419d",
                               "cluster3" = "#238b45"))+
  theme(axis.text.x = element_text(vjust = 1, hjust = 1, angle = 30)) +
  theme(legend.position = "top")
dev.off()
```

# Analysis: Cluster & CMS analysis
## calculate CMS subtype
```{r}
phe = geo$phe
emat = geo$exp
CMS.res = CMScaller::CMScaller(emat, templates = CMScaller::templates.CMS,
                    rowNames = "symbol")
saveRDS(CMS.res, file = "01.datatidy/geoCMS.rds")
```

## Visualization
```{r}
cluster = readRDS("04.rmps_cluster/cluster.rds")
geo = readRDS("01.datatidy/GEO_afterBatch.rds")
phe = geo$phe
cluster = cluster[[3]]$consensusClass

CMS.res = readRDS("01.datatidy/geoCMS.rds")
dat = data.frame(CMS = CMS.res$prediction, 
                 cluster = as.character(cluster))
dat = na.omit(dat)

dat$Cluster = paste("cluster", dat$cluster, sep = "")
dat = data.frame(table(dat))
dat = split(dat, dat$CMS)
dat = lapply(dat, function(x){x$Percent = (x$Freq/sum(x$Freq))*100
return(x)})
dat = do.call(rbind, dat)
dat$Percent = round(dat$Percent, 2)
dat$CMS = factor(dat$CMS, levels = c("CMS1", "CMS2", "CMS3", "CMS4"))

pdf(file = "04.rmps_cluster/rmps_CMS.pdf", width = 2, height = 3.5)
ggplot(dat, aes(x = CMS, y = Percent, fill=Cluster)) +
  geom_bar(position='stack', stat='identity') + 
  theme_classic() + 
  theme(panel.grid=element_blank()) +
  ylab("Precent(100%)") +
  scale_fill_manual(values = c("cluster1" = "#ec7014",
                               "cluster2" = "#88419d",
                               "cluster3" = "#238b45"))+
  theme(axis.text.x = element_text(vjust = 1, hjust = 1, angle = 30)) +
  theme(legend.position = "top")
dev.off()
```
# Analysis: Cluster & TME
## calculate TME signature → ssGSEA
```{r}
geo = readRDS("01.datatidy/GEO_afterBatch.rds")
exprMatrix = geo$exp
sets = data.table::fread("04.rmps_cluster/TME.txt")
sets = split(sets$Symbol, sets$Signature)
gsva_TME <- GSVA::gsva(as.matrix(exprMatrix), sets,method='ssgsea',kcdf='Gaussian',abs.ranking=TRUE)
saveRDS(gsva_TME, file = "04.rmps_cluster/geo_gsvaTME.rds")
```

## DE analysis
```{r}
library(limma)
gsva_TME = readRDS("04.rmps_cluster/geo_gsvaTME.rds")
clusters = readRDS("04.rmps_cluster/cluster.rds")
clusters = clusters[[3]]$consensusClass

LimmaDeg = function(TreatDat, ConDat){
  list <- c(rep("Treat", ncol(TreatDat)), 
            rep("Control", ncol(ConDat))) %>% factor(., levels = c("Control", "Treat"), ordered = F)
  list <- model.matrix(~factor(list)+0)  #把group设置成一个model matrix
  colnames(list) <- c("Control", "Treat")
  df = cbind(TreatDat, ConDat) 
  # df = apply(df, 2, as.numeric)
  df.fit <- lmFit(df, list)  ## 数据与list进行匹配
  
  df.matrix <- makeContrasts(Treat - Control, levels = list)
  fit <- contrasts.fit(df.fit, df.matrix)
  fit <- eBayes(fit)
  tempOutput <- topTable(fit,n = Inf, adjust = "fdr")
  return(tempOutput)
}

cluster1 = gsva_TME[,clusters == 1]
cluster2 = gsva_TME[,clusters == 2]
cluster3 = gsva_TME[,clusters == 3]

cluster1_vs_cluster2 = LimmaDeg(cluster1, cluster2)
cluster1_vs_cluster2$Type = "cluster1_vs_cluster2"
cluster2_vs_cluster3 = LimmaDeg(cluster2, cluster3)
cluster2_vs_cluster3$Type = "cluster2_vs_cluster3"
cluster1_vs_cluster3 = LimmaDeg(cluster1, cluster3)
cluster1_vs_cluster3$Type = "cluster1_vs_cluster3"
out = rbind(cluster1_vs_cluster2, cluster2_vs_cluster3,
            cluster1_vs_cluster3)
out$signature = c(row.names(cluster1_vs_cluster2),
                  row.names(cluster2_vs_cluster3),
                  row.names(cluster1_vs_cluster3))
write.table(out, file = "04.rmps_cluster/DETME_cluster.xls", sep = "\t", row.names = F)
```

```{r}
cluster1_up_1 = row.names(subset(cluster1_vs_cluster2, adj.P.Val<0.05 & logFC>0))
cluster1_up_2 = row.names(subset(cluster1_vs_cluster3, adj.P.Val<0.05 & logFC>0))
cluster1_up = intersect(cluster1_up_1, cluster1_up_2)

cluster2_up_1 = row.names(subset(cluster1_vs_cluster2, adj.P.Val<0.05 & logFC<0))
cluster2_up_2 = row.names(subset(cluster2_vs_cluster3, adj.P.Val<0.05 & logFC>0))
cluster2_up = intersect(cluster2_up_1, cluster2_up_2)

cluster3_up_1 = row.names(subset(cluster1_vs_cluster3, adj.P.Val<0.05 & logFC<0))
cluster3_up_2 = row.names(subset(cluster2_vs_cluster3, adj.P.Val<0.05 & logFC<0))
cluster3_up = intersect(cluster3_up_1, cluster3_up_2)
```

## visualization
```{r}
gsva_TME = readRDS("04.rmps_cluster/geo_gsvaTME.rds")
clusters = readRDS("04.rmps_cluster/cluster.rds")
dat = gsva_TME%>% data.frame()

scale_data = function(x){
  # x = (x-mean(x))/sd(x)
  # x = (x-min(x))/(max(x) - min(x))
  x = x - mean(x)  
  x = x / max(abs(x))
  return(x)
}
dat = apply(dat, 1, scale_data)
dat = t(dat)

Cluster = readRDS("04.rmps_cluster/cluster.rds")
geo = readRDS("01.datatidy/GEO_afterBatch.rds")
phe = geo$phe
sets = data.table::fread("04.rmps_cluster/TME.txt")

columnAnn = data.frame(id = colnames(dat), 
                       clusters = as.character(Cluster[[3]]$consensusClass),
                       datasets = phe$GSE[match(colnames(dat), phe$GSM)])
# columnAnn = columnAnn[order(columnAnn$clusters),]
rowAnn = data.frame(Sig = row.names(dat), 
                    type = sets$Function[match(row.names(dat), sets$Signature)])

library(ComplexHeatmap)
library(RColorBrewer)

# define colors
# col_fun = c("#756bb1", "#fd8d3c", "#de2d26")
col_fun = brewer.pal(5, "PuOr")
col_fun = colorRampPalette(rev(col_fun))(100)
# col_fun = colorRampPalette(col_fun)(100)

datasets_col = colorRampPalette(brewer.pal(10, "Paired"))(12)
names(datasets_col) = unique(columnAnn$datasets)

# annotation
column_ha = HeatmapAnnotation(Clusters = columnAnn$clusters,
                              Datasets = columnAnn$datasets,
                              col = list(Clusters = c("1" = "#ec7014", 
                                                      "2" = "#88419d", 
                                                      "3" = "#238b45"),
                                         Datasets = datasets_col))

row_ha = rowAnnotation(Type = rowAnn$type,
                       col = list(
                       Type = c("Angiogenesis_Fibrosis" = "#ec7014",
                                "Anti_Tumor_Micorenvironment" = "#88419d",
                                "Malignant_Cell_Properties" = "#238b45",
                                "Pro_Tumor_Microenvironment" = "#e7298a")
                       ))


dat = dat[rowAnn$Sig,columnAnn$id]

rowPath = split(rowAnn$Sig, rowAnn$type)

rowType = rowAnn$type
  
pdf(file = "04.rmps_cluster/rmps_TME_heatmap.pdf", width = 10, height = 5.5)
Heatmap(dat, col = col_fun,
        clustering_distance_columns  = "pearson",
        clustering_distance_rows  = "pearson",
        show_column_dend = FALSE,
        show_row_dend = FALSE,
        # column_order = columnAnn$id,
        show_column_names  = FALSE,
        show_row_names = TRUE,
        # row_names_gp = gpar(fontface = "italic"),
        top_annotation = column_ha,
        column_split = columnAnn$clusters,
        row_split = rowType,
        left_annotation = row_ha,
        # right_annotation = 
        # rowAnnotation(textbox = anno_textbox(rowType, rowPath,
        #                                      gp = gpar(fontsize = 12)))
        )
dev.off()
```

# Analysis: Cluster & Cibersort analysis
## Calculate → cibersort
```{r}
source("04.rmps_cluster/CIBERSORT.R")
geo = readRDS("01.datatidy/GEO_afterBatch.rds")
Exp = geo$exp
results = CIBERSORT("04.rmps_cluster/ref.txt", Exp, perm=100, QN=TRUE)
saveRDS(results, file = "04.rmps_cluster/geo_cibersort.rds")
```

## DE analysis
```{r}
library(limma)
results = readRDS("04.rmps_cluster/geo_cibersort.rds")
results = t(results[,1:22])
clusters = readRDS("04.rmps_cluster/cluster.rds")
clusters = clusters[[3]]$consensusClass

LimmaDeg = function(TreatDat, ConDat){
  list <- c(rep("Treat", ncol(TreatDat)), 
            rep("Control", ncol(ConDat))) %>% factor(., levels = c("Control", "Treat"), ordered = F)
  list <- model.matrix(~factor(list)+0)  #把group设置成一个model matrix
  colnames(list) <- c("Control", "Treat")
  df = cbind(TreatDat, ConDat) 
  # df = apply(df, 2, as.numeric)
  df.fit <- lmFit(df, list)  ## 数据与list进行匹配
  
  df.matrix <- makeContrasts(Treat - Control, levels = list)
  fit <- contrasts.fit(df.fit, df.matrix)
  fit <- eBayes(fit)
  tempOutput <- topTable(fit,n = Inf, adjust = "fdr")
  return(tempOutput)
}

cluster1 = results[,clusters == 1]
cluster2 = results[,clusters == 2]
cluster3 = results[,clusters == 3]

cluster1_vs_cluster2 = LimmaDeg(cluster1, cluster2)
cluster1_vs_cluster2$Type = "cluster1_vs_cluster2"
cluster2_vs_cluster3 = LimmaDeg(cluster2, cluster3)
cluster2_vs_cluster3$Type = "cluster2_vs_cluster3"
cluster1_vs_cluster3 = LimmaDeg(cluster1, cluster3)
cluster1_vs_cluster3$Type = "cluster1_vs_cluster3"
out = rbind(cluster1_vs_cluster2, cluster2_vs_cluster3,
            cluster1_vs_cluster3)
out$signature = c(row.names(cluster1_vs_cluster2),
                  row.names(cluster2_vs_cluster3),
                  row.names(cluster1_vs_cluster3))
write.table(out, file = "04.rmps_cluster/DEComponent_cluster.xls", sep = "\t", row.names = F)
```

## visualization
```{r}
library(dplyr)
results = readRDS("04.rmps_cluster/geo_cibersort.rds")
clusters = readRDS("04.rmps_cluster/cluster.rds")
dat = results[,1:22] %>% data.frame() %>%
  mutate(cluster = paste0("cluster", clusters[[3]]$consensusClass, sep = "")) 
dat = reshape2::melt(dat)

colnames(dat) = c("Group", "Cell", "value")
p = ggplot(dat, aes(x = Cell, y = value, color = Group)) +
  geom_boxplot(aes(x = Cell), width = 0.8, outlier.shape = NA) +
  # geom_jitter(aes(x = Gene, color = Type), size = 0.1, alpha = 0.3,width = 0.35) +
  geom_point(position=position_jitterdodge(jitter.width = 0.2, 
                                           dodge.width = 0.8), 
             size = 0.1, alpha = 0.1)+
  theme_classic()+
  theme(panel.grid=element_blank()) +
  theme(axis.text.x = element_text(vjust = 1, hjust = 1, angle = 30)) +
  ylab("Immune Component") +
  scale_color_manual(values = c("cluster1" = "#ec7014", 
                               "cluster2" = "#88419d", 
                               "cluster3" = "#238b45")) +  
  stat_compare_means(label = "p.signif") +
  theme(legend.position = "top") +
  theme(axis.title.x = element_blank(),
        axis.text = element_text(colour = "black"))
p
ggsave(p, filename = "04.rmps_cluster/rmps_clusterCibersort.pdf", width = 8, height = 4)
```

# Analysis: Cluster & Infiltration analysis
## Calculate → ssGSEA
```{r}
geo = readRDS("01.datatidy/GEO_afterBatch.rds")
exprMatrix = geo$exp
sets = data.table::fread("04.rmps_cluster/ImmCell.txt")
sets = split(sets$Gene, sets$CellType)
gsva_cell <- GSVA::gsva(as.matrix(exprMatrix), sets,method='ssgsea',kcdf='Gaussian',abs.ranking=TRUE)
saveRDS(gsva_cell, file = "04.rmps_cluster/geo_gsvaCell.rds")
```
## DE analysis
```{r}
library(limma)
gsva_cell = readRDS("04.rmps_cluster/geo_gsvaCell.rds")
clusters = readRDS("04.rmps_cluster/cluster.rds")
clusters = clusters[[3]]$consensusClass

LimmaDeg = function(TreatDat, ConDat){
  list <- c(rep("Treat", ncol(TreatDat)), 
            rep("Control", ncol(ConDat))) %>% factor(., levels = c("Control", "Treat"), ordered = F)
  list <- model.matrix(~factor(list)+0)  #把group设置成一个model matrix
  colnames(list) <- c("Control", "Treat")
  df = cbind(TreatDat, ConDat) 
  # df = apply(df, 2, as.numeric)
  df.fit <- lmFit(df, list)  ## 数据与list进行匹配
  
  df.matrix <- makeContrasts(Treat - Control, levels = list)
  fit <- contrasts.fit(df.fit, df.matrix)
  fit <- eBayes(fit)
  tempOutput <- topTable(fit,n = Inf, adjust = "fdr")
  return(tempOutput)
}

cluster1 = gsva_cell[,clusters == 1]
cluster2 = gsva_cell[,clusters == 2]
cluster3 = gsva_cell[,clusters == 3]

cluster1_vs_cluster2 = LimmaDeg(cluster1, cluster2)
cluster1_vs_cluster2$Type = "cluster1_vs_cluster2"
cluster2_vs_cluster3 = LimmaDeg(cluster2, cluster3)
cluster2_vs_cluster3$Type = "cluster2_vs_cluster3"
cluster1_vs_cluster3 = LimmaDeg(cluster1, cluster3)
cluster1_vs_cluster3$Type = "cluster1_vs_cluster3"
out = rbind(cluster1_vs_cluster2, cluster2_vs_cluster3,
            cluster1_vs_cluster3)
out$signature = c(row.names(cluster1_vs_cluster2),
                  row.names(cluster2_vs_cluster3),
                  row.names(cluster1_vs_cluster3))
write.table(out, file = "04.rmps_cluster/DEInfiltration_cluster.xls", sep = "\t", row.names = F)
```
## visualization
```{r}
gsva_cell = readRDS("04.rmps_cluster/geo_gsvaCell.rds")
clusters = readRDS("04.rmps_cluster/cluster.rds")
dat = gsva_cell%>% t() %>% data.frame() %>%
  mutate(cluster = paste0("cluster", clusters[[3]]$consensusClass, sep = "")) 
dat = reshape2::melt(dat)

colnames(dat) = c("Group", "Cell", "value")
p = ggplot(dat, aes(x = Cell, y = value, color = Group)) +
  geom_boxplot(aes(x = Cell), width = 0.8, outlier.shape = NA) +
  # geom_jitter(aes(x = Gene, color = Type), size = 0.1, alpha = 0.3,width = 0.35) +
  geom_point(position=position_jitterdodge(jitter.width = 0.2, 
                                           dodge.width = 0.8), 
             size = 0.1, alpha = 0.1)+
  theme_classic()+
  theme(panel.grid=element_blank()) +
  theme(axis.text.x = element_text(vjust = 1, hjust = 1, angle = 30)) +
  ylab("log2(FPKM+1)") +
  scale_color_manual(values = c("cluster1" = "#ec7014", 
                               "cluster2" = "#88419d", 
                               "cluster3" = "#238b45")) +  
  stat_compare_means(label = "p.signif") +
  theme(legend.position = "top")+
  theme(axis.title.x = element_blank(),
        axis.text = element_text(colour = "black"))

ggsave(p, filename = "04.rmps_cluster/geo_clusterCell.pdf", width = 8, height = 4)
```

# Analysis: Cluster & EMT analysis
# Analysis: EMT signature
## load data → ssGSEA
```{r}
geo = readRDS("01.datatidy/GEO_afterBatch.rds")
exprMatrix = geo$exp
sets = data.table::fread("04.rmps_cluster/EMT.txt")
sets = split(sets$Symbol, sets$Signature)
gsva_EMT <- GSVA::gsva(as.matrix(exprMatrix), sets,method='ssgsea',kcdf='Gaussian',abs.ranking=TRUE)
saveRDS(gsva_EMT, file = "04.rmps_cluster/geo_gsvaEMT.rds")
```
## DE analysis
```{r}
library(limma)
gsva_EMT = readRDS("04.rmps_cluster/geo_gsvaEMT.rds")
clusters = readRDS("04.rmps_cluster/cluster.rds")
clusters = clusters[[3]]$consensusClass

LimmaDeg = function(TreatDat, ConDat){
  list <- c(rep("Treat", ncol(TreatDat)), 
            rep("Control", ncol(ConDat))) %>% factor(., levels = c("Control", "Treat"), ordered = F)
  list <- model.matrix(~factor(list)+0)  #把group设置成一个model matrix
  colnames(list) <- c("Control", "Treat")
  df = cbind(TreatDat, ConDat) 
  # df = apply(df, 2, as.numeric)
  df.fit <- lmFit(df, list)  ## 数据与list进行匹配
  
  df.matrix <- makeContrasts(Treat - Control, levels = list)
  fit <- contrasts.fit(df.fit, df.matrix)
  fit <- eBayes(fit)
  tempOutput <- topTable(fit,n = Inf, adjust = "fdr")
  return(tempOutput)
}

cluster1 = gsva_EMT[,clusters == 1]
cluster2 = gsva_EMT[,clusters == 2]
cluster3 = gsva_EMT[,clusters == 3]

cluster1_vs_cluster2 = LimmaDeg(cluster1, cluster2)
cluster1_vs_cluster2$Type = "cluster1_vs_cluster2"
cluster2_vs_cluster3 = LimmaDeg(cluster2, cluster3)
cluster2_vs_cluster3$Type = "cluster2_vs_cluster3"
cluster1_vs_cluster3 = LimmaDeg(cluster1, cluster3)
cluster1_vs_cluster3$Type = "cluster1_vs_cluster3"
out = rbind(cluster1_vs_cluster2, cluster2_vs_cluster3,
            cluster1_vs_cluster3)
out$signature = c(row.names(cluster1_vs_cluster2),
                  row.names(cluster2_vs_cluster3),
                  row.names(cluster1_vs_cluster3))
write.table(out, file = "04.rmps_cluster/DEEMT_cluster.xls", sep = "\t", row.names = F)
```
## visualization
```{r}
gsva_EMT = readRDS("04.rmps_cluster/geo_gsvaEMT.rds")
clusters = readRDS("04.rmps_cluster/cluster.rds")
dat = gsva_EMT%>% t() %>% data.frame() %>%
  mutate(cluster = paste0("cluster", clusters[[3]]$consensusClass, sep = "")) 
dat = reshape2::melt(dat)

colnames(dat) = c("Group", "Sig", "value")
# dat = subset(dat, Sig %in% c("Angiogenesis","EMT1", "EMT2", "EMT3",
#                              "CD8.T.effector"))
p = ggplot(dat, aes(x = Sig, y = value, color = Group)) +
  geom_boxplot(aes(x = Sig), width = 0.8, outlier.shape = NA) +
  # geom_jitter(aes(x = Gene, color = Type), size = 0.1, alpha = 0.3,width = 0.35) +
  geom_point(position=position_jitterdodge(jitter.width = 0.2, 
                                           dodge.width = 0.8), 
             size = 0.1, alpha = 0.1)+
  theme_classic()+
  theme(panel.grid=element_blank()) +
  theme(axis.text.x = element_text(vjust = 1, hjust = 1, angle = 30)) +
  ylab("Enrichment Score") +
  scale_color_manual(values = c("cluster1" = "#ec7014", 
                               "cluster2" = "#88419d", 
                               "cluster3" = "#238b45")) +  
  stat_compare_means(label = "p.signif") +
  theme(legend.position = "top")+
  theme(axis.title.x = element_blank(),
        axis.text = element_text(colour = "black"))
# p
ggsave(p, filename = "04.rmps_cluster/geo_EMT.pdf", width = 6, height = 4)
```