# Calculate Enrichment Score
```{r}
dir.create("05.rmps_scSeq")
sc_MMR = readRDS("01.datatidy/sc_MMR.rds")
rmps = readRDS("01.datatidy/rmps.rds")
sc_Ann = data.table::fread("01.datatidy/scAnn.txt", header = F)
colnames(sc_Ann) = c("symbol", "ensg")
sc_Ann$ensg = stringr::str_split(sc_Ann$ensg, "_", simplify = T)[,1]

## subset RMP scSeq object
convertRownames <- function(seu_object, scAnn) {
  new_rownames <- sc_Ann$symbol[match(rownames(seu_object), sc_Ann$ensg)]
  rownames(seu_object@assays$RNA@counts) <- new_rownames
  rownames(seu_object@assays$RNA@data) <- new_rownames
  return(seu_object)
}
sc_MMR = convertRownames(seu_object = sc_MMR, scAnn = sc_Ann)

table(rmps$Symbol %in% row.names(sc_MMR))
scRMP =sc_MMR[rmps$Symbol,]
saveRDS(scRMP, file = "01.datatidy/scRMP.rds")

## calculate rmp enrichment score
library(Seurat)
steps = seq(from = 1, to = 371223, by = 1000)
steps[373] = 371223

rmp_enrich = list()
for (i in 1:(length(steps)-1)) {
  tempdat = sc_MMR@assays$RNA@counts[,steps[i]:steps[i+1]] # 371223
  sets = list(rmps = rmps$Symbol)
  gsva_RMPS <- GSVA::gsva(tempdat,
                          sets,method='ssgsea',kcdf='Gaussian',abs.ranking=TRUE)
  rmp_enrich[[i]] = gsva_RMPS
}
saveRDS(rmp_enrich, file = "05.scSeq/rmp_enrich.rds")

## get meta information
metaIn = sc_MMR@meta.data
tsne = data.table::fread("05.rmps_scSeq/tsne.tsv")
for (i in 1:length(rmp_enrich)) {
  if (i == 1) {
    rmp_enrich[[i]] = rmp_enrich[[i]]@x
  }else{
    rmp_enrich[[i]] = rmp_enrich[[i]]@x[2:length( rmp_enrich[[i]]@x)]
  }
}
rmps_score = unlist(rmp_enrich)
metaIn$rmps_score = rmps_score
metaIn$sampleID = row.names(metaIn)
metaIn = dplyr::left_join(metaIn, tsne, by = "sampleID")
saveRDS(metaIn, file = "05.rmps_scSeq/rmp_enrich_meta.rds")
```

# Analysis: Differential RMPS_enrich: Tumor vs Normal
## scatter plot: landscape
```{r}
library(ggplot2)
library(RColorBrewer)
library(ggpubr)
getPalette = colorRampPalette(brewer.pal(8, "Set1"))(8)
colorGrad = c("#e66101",
               "#fdb863",
               "#f7f7f7",
               "#b2abd2",
               "#5e3c99")
dat = readRDS("05.rmps_scSeq/rmp_enrich_meta.rds")

p1 = dat %>% subset(SPECIMEN_TYPE == "N") %>%
  ggplot(., aes(x = ydata_1, y = ydata_2, color = clTopLevel)) +
  geom_point() + 
  theme_classic() +
  scale_color_manual(values = rev(getPalette)) + 
  theme(axis.line = element_blank(), axis.title = element_blank(), 
        axis.text = element_blank(), axis.ticks = element_blank())
p1
p2 = dat %>% subset(SPECIMEN_TYPE == "T") %>%
  ggplot(., aes(x = ydata_1, y = ydata_2, color = clTopLevel)) +
  geom_point() + 
  theme_classic() +
  scale_color_manual(values = rev(getPalette)) + 
  theme(axis.line = element_blank(), axis.title = element_blank(), 
        axis.text = element_blank(), axis.ticks = element_blank())

p3 = dat %>% subset(SPECIMEN_TYPE == "N") %>%
  ggplot(., aes(x = ydata_1, y = ydata_2, color=rmps_score)) +
  geom_point() + 
  theme_classic() +
  scale_colour_gradientn(colors=rev(colorGrad))+ 
  theme(axis.line = element_blank(), axis.title = element_blank(), 
        axis.text = element_blank(), axis.ticks = element_blank())

p4 = dat %>% subset(SPECIMEN_TYPE == "T") %>%
  ggplot(., aes(x = ydata_1, y = ydata_2, color=rmps_score)) +
  geom_point() + 
  theme_classic() +
  scale_colour_gradientn(colors=rev(colorGrad)) + 
  theme(axis.line = element_blank(), axis.title = element_blank(), 
        axis.text = element_blank(), axis.ticks = element_blank())
p = (p1 | p2 ) / (p3 | p4)
# ggsave(p1, filename = "05.scSeq/sc_normal.pdf", width = 5, height = 4)
# ggsave(p2, filename = "05.scSeq/sc_tumor.pdf", width = 5, height = 4)
# ggsave(p3, filename = "05.scSeq/sc_normal_rmps.pdf", width = 5, height = 4)
# ggsave(p4, filename = "05.scSeq/sc_tumor_rmps.pdf", width = 5, height = 4)
ggsave(p, filename = "05.rmps_scSeq/sc_enrichment.png", width = 10, height = 8, dpi = 2048)

```

## Analysis: rmps_enrich_main_type
```{r}
dat = readRDS("05.rmps_scSeq/rmp_enrich_meta.rds")
dat = split(dat, dat$SPECIMEN_TYPE)

Analysis = function(obj){
  obj = split(obj, obj$clTopLevel)
  mean_value = lapply(obj, function(x){x = mean(x$rmps_score)})
  sd_value = lapply(obj, function(x){x = sd(x$rmps_score)})
  n_sum = lapply(obj, function(x){x = nrow(x)})
  n_sum_total = sum(unlist(n_sum))
  component = lapply(n_sum, function(x){x = x/n_sum_total*100})
  out = data.frame(cell = names(obj), 
                   mean_rmps = round(unlist(mean_value),3),
                   sd = paste("±", round(unlist(sd_value),3), sep = ""),
                   component = round(unlist(component),3))
  return(out)
}

res = lapply(dat, Analysis)
res = do.call(rbind, res)
res$Type = stringr::str_split(row.names(res), "[.]", simplify = T)[,1]
write.table(res, file = "05.rmps_scSeq/enrich_landscape_tumor_normal.xls",
      row.names = F, sep = "\t")
saveRDS(res, file = "05.rmps_scSeq/summ_land_N_T.rds")
```

## violin plot: landscape
```{r}
library(RColorBrewer)
library(dplyr)
library(ggplot2)
dat = readRDS("05.rmps_scSeq/rmp_enrich_meta.rds")
summ = readRDS("05.rmps_scSeq/summ_land_N_T.rds")
getPalette = colorRampPalette(brewer.pal(8, "Set1"))(11)
# input = dat %>% subset(SPECIMEN_TYPE == "T")
# input = dat 
# %>% subset(SPECIMEN_TYPE == "N")
N_summ = subset(summ, Type == "N")

df = subset(dat, SPECIMEN_TYPE == "N")
p1 = df %>% 
  ggplot(mapping = aes(x = clTopLevel, y = rmps_score)) + 
  geom_violin(aes(fill = clTopLevel)) +
  geom_point(data = N_summ,
             mapping = aes(x = cell, y = mean_rmps),
             color="black") +
  geom_line(data = N_summ, 
            mapping = aes(x = cell, y = mean_rmps),
            color="black", group=1)+
  # geom_text(data = N_summ, mapping = aes(x = cell, label = component)) +
  theme_classic() +
  ylab("Enrichment score") +
  theme(axis.title.x = element_blank()) +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(vjust = 1, hjust = 1, angle = 45))+
  scale_fill_manual(values = rev(getPalette))


T_summ = subset(summ, Type == "T")
df = subset(dat, SPECIMEN_TYPE == "T")
p2 = df %>% 
  ggplot(mapping = aes(x = clTopLevel, y = rmps_score)) + 
  geom_violin(aes(fill = clTopLevel)) +
  geom_point(data = T_summ,
             mapping = aes(x = cell, y = mean_rmps),
             color="black") +
  geom_line(data = T_summ, 
            mapping = aes(x = cell, y = mean_rmps),
            color="black", group=1)+
  theme_classic() +
  ylab("Enrichment score") +
  theme(axis.title.x = element_blank()) +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(vjust = 1, hjust = 1, angle = 45))+
  scale_fill_manual(values = rev(getPalette))

p = (p1 / p2)
ggsave(p, filename = "05.rmps_scSeq/sc_enrichment_violin.pdf", width = 5, height = 4)
```

## scatter plot: epi
```{r}
dat = readRDS("05.rmps_scSeq/rmp_enrich_meta.rds")
getPalette = colorRampPalette(brewer.pal(8, "Set1"))(11)
p1 = dat %>% subset(clTopLevel == "Epi") %>%
  subset(SPECIMEN_TYPE == "N") %>%
  ggplot(., aes(x = ydata_1, y = ydata_2, color=clFull)) +
  geom_point() + 
  theme_classic() +
  scale_color_manual(values = rev(getPalette)) + 
  # scale_colour_gradientn(colors=rev(colorGrad)) + 
  theme(axis.line = element_blank(), axis.title = element_blank(), 
        axis.text = element_blank(), axis.ticks = element_blank())

p2 = dat %>% subset(clTopLevel == "Epi") %>%
  subset(SPECIMEN_TYPE == "T") %>%
  ggplot(., aes(x = ydata_1, y = ydata_2, color=clFull)) +
  geom_point() + 
  theme_classic() +
  scale_color_manual(values = rev(getPalette)) + 
  # scale_colour_gradientn(colors=rev(colorGrad)) + 
  theme(axis.line = element_blank(), axis.title = element_blank(), 
        axis.text = element_blank(), axis.ticks = element_blank())


p3 = dat %>% subset(clTopLevel == "Epi") %>% subset(SPECIMEN_TYPE == "N") %>%
  ggplot(., aes(x = ydata_1, y = ydata_2, color=rmps_score)) +
  geom_point() + 
  theme_classic() +
  scale_colour_gradientn(colors=rev(colorGrad))+ 
  theme(axis.line = element_blank(), axis.title = element_blank(), 
        axis.text = element_blank(), axis.ticks = element_blank())

p4 = dat %>% subset(clTopLevel == "Epi")%>% subset(SPECIMEN_TYPE == "T") %>%
  ggplot(., aes(x = ydata_1, y = ydata_2, color=rmps_score)) +
  geom_point() + 
  theme_classic() +
  scale_colour_gradientn(colors=rev(colorGrad)) + 
  theme(axis.line = element_blank(), axis.title = element_blank(), 
        axis.text = element_blank(), axis.ticks = element_blank())

p = (p1 | p2 ) / (p3 | p4)
ggsave(p, filename = "05.rmps_scSeq/sc_enrichment_Epi.png", width = 10, height = 8, dpi = 2048)
ggsave(p1, filename = "05.rmps_scSeq/sc_normal_Epi.pdf", width = 5, height = 4)
ggsave(p2, filename = "05.rmps_scSeq/sc_tumor_Epi.pdf", width = 5, height = 4)
ggsave(p3, filename = "05.rmps_scSeq/sc_normal_rmps_Epi.pdf", width = 5, height = 4)
ggsave(p4, filename = "05.rmps_scSeq/sc_tumor_rmps_Epi.pdf", width = 5, height = 4)
```


## Analysis: rmps_enrich_epi_type
```{r}
dat = readRDS("05.rmps_scSeq/rmp_enrich_meta.rds")
dat = subset(dat, clTopLevel == "Epi")
dat = split(dat, dat$SPECIMEN_TYPE)

Analysis = function(obj){
  obj = split(obj, obj$clFull)
  mean_value = lapply(obj, function(x){x = mean(x$rmps_score)})
  sd_value = lapply(obj, function(x){x = sd(x$rmps_score)})
  n_sum = lapply(obj, function(x){x = nrow(x)})
  n_sum_total = sum(unlist(n_sum))
  component = lapply(n_sum, function(x){x = x/n_sum_total*100})
  out = data.frame(cell = names(obj), 
                   mean_rmps = round(unlist(mean_value),3),
                   sd = paste("±", round(unlist(sd_value),3), sep = ""),
                   component = round(unlist(component),3))
  return(out)
}

res = lapply(dat, Analysis)
res = do.call(rbind, res)
res$Type = stringr::str_split(row.names(res), "[.]", simplify = T)[,1]
write.table(res, file = "05.rmps_scSeq/enrich_epi_tumor_normal.xls",
      row.names = F, sep = "\t")
saveRDS(res, file = "05.rmps_scSeq/summ_epi_N_T.rds")
```

## Analysis: rmps_enrich_other_type
"Epi"     "TNKILC"  "Strom"   "Myeloid" "B"       "Mast"    "Plasma"
```{r}
type = "Plasma" 
dat = readRDS("05.rmps_scSeq/rmp_enrich_meta.rds")
dat = subset(dat, clTopLevel == type)
dat = split(dat, dat$SPECIMEN_TYPE)

Analysis = function(obj){
  obj = split(obj, obj$clFull)
  mean_value = lapply(obj, function(x){x = mean(x$rmps_score)})
  sd_value = lapply(obj, function(x){x = sd(x$rmps_score)})
  n_sum = lapply(obj, function(x){x = nrow(x)})
  n_sum_total = sum(unlist(n_sum))
  component = lapply(n_sum, function(x){x = x/n_sum_total*100})
  out = data.frame(cell = names(obj), 
                   mean_rmps = round(unlist(mean_value),3),
                   sd = paste("±", round(unlist(sd_value),3), sep = ""),
                   component = round(unlist(component),3))
  return(out)
}

res = lapply(dat, Analysis)
res = do.call(rbind, res)
res$Type = stringr::str_split(row.names(res), "[.]", simplify = T)[,1]
write.table(res, file = paste("05.rmps_scSeq/enrich", type, "tumor_normal.xls", sep = "_"),
      row.names = F, sep = "\t")
saveRDS(res, file = paste("05.rmps_scSeq/summ", type,"N_T.rds", sep = ""))
```

## violin plot: Epi
```{r}
getPalette = colorRampPalette(brewer.pal(8, "Set1"))(11)
dat = readRDS("05.rmps_scSeq/rmp_enrich_meta.rds")
summ = readRDS("05.rmps_scSeq/summ_epi_N_T.rds")
df = subset(dat, SPECIMEN_TYPE == "N" & clTopLevel == "Epi")
N_summ = subset(summ, Type == "N")
p1 = df %>%
  ggplot(., aes(x = clFull, y = rmps_score)) + 
  geom_violin(aes(fill = clFull)) +
  geom_point(data = N_summ,
             mapping = aes(x = cell, y = mean_rmps),
             color="black") +
  geom_line(data = N_summ, 
            mapping = aes(x = cell, y = mean_rmps),
            color="black", group=1)+
  theme_classic() +
  ylab("Enrichment score") +
  theme(axis.title.x = element_blank()) +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(vjust = 1, hjust = 1, angle = 45))+
  scale_fill_manual(values = rev(getPalette)) 

T_summ = subset(summ, Type == "T")
df = subset(dat, SPECIMEN_TYPE == "T" & clTopLevel == "Epi")
p2 = df %>% 
  ggplot(., aes(x = clFull, y = rmps_score)) +
  geom_violin(aes(fill = clFull)) +
  geom_point(data = N_summ,
             mapping = aes(x = cell, y = mean_rmps),
             color="black") +
  geom_line(data = N_summ, 
            mapping = aes(x = cell, y = mean_rmps),
            color="black", group=1)+  theme_classic() +
  ylab("Enrichment score") +
  theme(axis.title.x = element_blank()) +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(vjust = 1, hjust = 1, angle = 45))+
  scale_fill_manual(values = rev(getPalette)) 
p = (p1 / p2)
ggsave(p, filename = "05.rmps_scSeq/sc_enrichment_violin_Epi.pdf", width = 5, height = 7)
```

# Analysis: Gene Set Enrichment: Normal Vs MMR Vs MMd
## kruskal.test
```{r}
library(ggpubr)
dat = readRDS("05.rmps_scSeq/rmp_enrich_meta.rds")
dat = split(dat, dat$clFull)

ReshapeDat = function(subdat){
  # subdat = dat$`cS25 (Fibro CCL8+)`
  subdat$type = subdat$MMRStatus
  subdat$type = ifelse(subdat$SPECIMEN_TYPE == "N", "Normal", subdat$type)
  kt.res = kruskal.test(rmps_score~type, data = subdat)
  mean_subdat = split(subdat, subdat$type)
  mean_subdat = lapply(mean_subdat, function(x){x = mean(x$rmps_score)})
  mean_subdat = do.call(cbind, mean_subdat)
  out = data.frame(mean_subdat)
  out$pval = kt.res$p.value
  return(out)
}

dat = lapply(dat, ReshapeDat)
for (i in 1:length(dat)) {
  if (dim(dat[[i]])[2]<4) {
    dat[[i]] = data.frame(MMRd = NA, MMRp = NA, Normal = NA, pval = NA)
  }
}
dat = do.call(rbind, dat)
dat = na.omit(dat)
dat$label = ifelse(dat$pval>=0.05, "NS", 
                   ifelse(dat$pval>=0.01, "*",
                    ifelse(dat$pval>=0.001, "**",
                    ifelse(dat$pval>=0.0001, "***", "****"))))
dat$cell = row.names(dat)
temp_Dat = dat
write.table(temp_Dat, file = "05.rmps_scSeq/Enrichment_Diff.xls", sep = "\t", row.names = F)
```


## visualiation:    kruskal.test
B     Epi    Mast Myeloid  Plasma   Strom  TNKILC
```{r}
getPalette = colorRampPalette(brewer.pal(8, "Set1"))(11)
cols = getPalette[6]
cell_hir = data.frame(hir1 = metaIn$clTopLevel,
                      hir2 = metaIn$clMidway,
                      hir3 = metaIn$clFull)
cell_hir = cell_hir[!duplicated(cell_hir$hir3),]
cell_hir = subset(cell_hir, hir1 == "Strom")
dat = subset(temp_Dat, cell %in% cell_hir$hir3)
dat = split(dat, dat$cell)

Draw = function(subdat){
  # subdat = dat$`cB1 (B IGD+IgM+)`
  input = reshape2::melt(subdat)[1:3,c(3,4)]
  colnames(input) = c("type", "value")
  input$type = factor(input$type, levels = c("Normal", "MMRp", "MMRd"))
  
  p = ggplot(input, aes(x=type, y=value, group=1)) + 
    geom_line(color = cols, linewidth = 2, alpha = 0.8) +
    geom_point(color = cols, size = 3, alpha = 0.8) +
    theme_classic() +
    ylab(subdat$cell[1])+
    theme(axis.title.x = element_blank()) +
    geom_text(label = subdat$label[1], size = 6) +
    theme(panel.grid=element_blank())+
     theme(axis.text.x = element_text(vjust = 1, hjust = 1, angle = 45))
  return(p)
}

p.res = lapply(dat, Draw)
p.out = patchwork::wrap_plots(p.res, 
                    nrow = floor(nrow(cell_hir)/8), ncol = 8)
ggsave(p.out, filename = "05.scSeq/Enricnment_Strom.pdf", width = 1.6*8, height = 1.8*4)
```

## Analysis: rmps_enrich_type
"Epi"     "TNKILC"  "Strom"   "Myeloid" "B"       "Mast"    "Plasma"
```{r}
type = "Plasma" 
dat = readRDS("05.rmps_scSeq/rmp_enrich_meta.rds")
dat = subset(dat, clTopLevel == type)
dat$type = dat$MMRStatus
dat$type = ifelse(dat$SPECIMEN_TYPE == "N", "Normal", dat$type)
dat = split(dat, dat$type)

Analysis = function(obj){
  obj = split(obj, obj$clFull)
  mean_value = lapply(obj, function(x){x = mean(x$rmps_score)})
  sd_value = lapply(obj, function(x){x = sd(x$rmps_score)})
  n_sum = lapply(obj, function(x){x = nrow(x)})
  n_sum_total = sum(unlist(n_sum))
  component = lapply(n_sum, function(x){x = x/n_sum_total*100})
  out = data.frame(cell = names(obj), 
                   mean_rmps = round(unlist(mean_value),3),
                   sd = paste("±", round(unlist(sd_value),3), sep = ""),
                   component = round(unlist(component),3))
  return(out)
}

res = lapply(dat, Analysis)
res = do.call(rbind, res)
res$Type = stringr::str_split(row.names(res), "[.]", simplify = T)[,1]
write.table(res, file = paste("05.rmps_scSeq/enrich", type, "N_T_MMR.xls", sep = "_"),
      row.names = F, sep = "\t")
saveRDS(res, file = paste("05.rmps_scSeq/summ", type,"N_T_MMR.rds", sep = ""))
```

# Analysis: DEGs:MMRd vs MMRp
```{r}
library(limma)
library(Seurat)
rmps_sc = readRDS("01.datatidy/scRMP.rds")
LimmaDeg = function(TreatDat, ConDat){
  list <- c(rep("Treat", ncol(TreatDat)), 
            rep("Control", ncol(ConDat))) %>% factor(., levels = c("Control", "Treat"), ordered = F)
  list <- model.matrix(~factor(list)+0)  #把group设置成一个model matrix
  colnames(list) <- c("Control", "Treat")
  df = cbind(TreatDat, ConDat) 
  # df = apply(df, 2, as.numeric)
  df.fit <- lmFit(df, list)  ## 数据与list进行匹配
  
  df.matrix <- makeContrasts(Treat - Control, levels = list)
  fit <- contrasts.fit(df.fit, df.matrix)
  fit <- eBayes(fit)
  tempOutput <- topTable(fit,n = Inf, adjust = "fdr")
  return(tempOutput)
}

metaIn = rmps_sc@meta.data
metaIn$type = ifelse(metaIn$SPECIMEN_TYPE == "N", "Normal", 
                      metaIn$MMRStatus)

metaIn = split(metaIn, metaIn$clFull)

res = list()
for (i in 1:length(metaIn)) {
  # dat = metaIn$`cB1 (B IGD+IgM+)`
  dat = metaIn[[i]]
  if ("MMRp" %in% unique(dat$type) & "MMRd" %in% unique(dat$type)) {
      normal_id = row.names(dat)[dat$type == "Normal"]
      MMRp_id = row.names(dat)[dat$type == "MMRp"]
      MMRd_id = row.names(dat)[dat$type == "MMRd"]
      datNormal = as.data.frame(rmps_sc@assays$RNA@counts[,normal_id])
      datMMRp = as.data.frame(rmps_sc@assays$RNA@counts[,MMRp_id])
      datMMRd = as.data.frame(rmps_sc@assays$RNA@counts[,MMRd_id])
      
      datMMRp_vs_datMMRd = LimmaDeg(datMMRp, datMMRd)
      datMMRp_vs_datMMRd$cell = names(metaIn)[i]
      datMMRp_vs_datMMRd$gene = row.names(datMMRp_vs_datMMRd)
      res[[i]] = datMMRp_vs_datMMRd
  }

}
res = do.call(rbind, res)
write.table(res, file = "05.rmps_scSeq/MMRp_MMRd.xls", sep = "\t", row.names = F)
saveRDS(res, file = "05.rmps_scSeq/DEG_cell_res.rds")
```

## visualization
```{r}
library(ggplot2)
library(RColorBrewer)
# colors = c("#bebada",
#            "#80b1d3",
#            "#8dd3c7",
#            # "#ffffb3",
#            "#fdb462",
#            "#fb8072")
colors = brewer.pal(10, "PuOr")

dat = subset(res, adj.P.Val<0.05 & abs(logFC)>0.2)
# dat = subset(res, adj.P.Val<0.05)
p = ggplot(dat, aes(x = cell, y = gene, size = -log10(adj.P.Val), color = logFC)) +
  geom_point() +
  theme_bw() +
  theme(axis.text.x = element_text(vjust = 1, hjust = 1, angle = 45)) +
  scale_colour_gradientn(colours = colorRampPalette(rev(colors))(100)) +
  theme()+
  theme(panel.grid=element_blank()) +
  theme(axis.text.y = element_text(face = "italic")) +
  theme(axis.title = element_blank()) 
# +
  # theme(legend.position = "bottom")
ggsave(p, filename = "05.rmps_scSeq/MMRd_MMRp_DERmps.pdf", width = 5, height = 8)
```
